---
editor_options: 
  chunk_output_type: console
---

# Exploring outputs from BirdNET  

Please note that thresholds were previously set for the target list of species, following analysis that were carried out here: 
https://vjjan91.github.io/paired-pointCount-acoustics/exploring-outputs-from-birdnet.html 

In this script, we use previously set thresholds for a subset of species in our study.  

## Install required libraries
```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(ggplot2)
library(data.table)
library(extrafont)
library(sf)
library(raster)
library(viridis) # for colorblind-friendly colors
library(gridExtra) # for arranging multiple plots

# for plotting
library(scales)
library(ggplot2)
library(ggspatial)
library(colorspace)
library(scico)
```

## Loading the acoustic data and filtering outputs above a threshold  

We are able to use previously set thresholds for all species, except the Gray-headed Canary-Flycatcher, Nilgiri Flycatcher or the Ashy Drongo. Excluding these three species, we created curated thresholds for each species above which the probability of detecting a species is at 0.95. For a few species for which the detector performed really well (proportion of true positives: false positives > 0.95), we set a random threshold value of 0.25 above which we expect the probability of species detected to be greater than or equal to 0.95. 
```{r}
## curate species thresholds
threshold <- read.csv("results/curatedThresholds-species-list.csv")

## load the BirdNET output file
birdnet_results <- read.csv("results/birdNET-outputs/BirdNET_CombinedTable.csv")

## filter species detections above the threshold we chose
birdnet_subset <- birdnet_results %>%
    inner_join(threshold, by = c("Common.name" = "species")) %>%
    filter(Confidence >= threshold) %>%
    mutate(filename_pattern = basename(File) %>% 
                        sub(".*\\\\([0-9]{8}_[0-9]{4}_[A-Z0-9]+)\\.WAV", "\\1", .)) %>%
  mutate(
    date = as.Date(str_extract(filename_pattern, "\\d{8}"), format = "%Y%m%d"),
    site_name = str_extract(filename_pattern, "^[A-Z]+"),
    site_protocol_code = paste0(site_name, "_MSF")
  )
```

## Including the Ashy Drongo

The model did not perform well enough for us to set thresholds for the Ashy Drongo and all instances were manually verified. So we manually validated every single detection (n=1281) to ascertain true and false positives. Please see this document for information on the validation procedure: https://zenodo.org/records/8415090  

```{r}
# load the .txt file containing true and false positives
ashy <- read.table("results/birdNET-segments/Ashy Drongo/ashy-drongo.selections.txt", sep = "\t", header = T)

## cleaning up the data 
ashy <- ashy %>%
  mutate(
    # extract confidence score
    confidence = str_extract(Begin.File, "^\\d+\\.\\d+"),
    
    # extract site_name
    site_name = str_extract(Begin.File, "[A-Z]+(?=_\\d{8})"),
    
    # extract date and format it
    date = as.Date(str_extract(Begin.File, "\\d{8}"), format = "%Y%m%d"),
    
    # create filename_pattern
    filename_pattern = str_extract(Begin.File, "[A-Z]+_\\d{8}_\\d{6}") %>% 
      paste0(".WAV"),
    
    # create site_protocol_code
    site_protocol_code = paste0(site_name, "_MSF")
  )

## filter only detections that are true positives and include scientific_name and common_name
ashy <- ashy %>%
  filter(Valid == 1) %>%
  mutate(scientific_name = "Dicrurus leucophaeus") %>%
  mutate(common_name = "Ashy Drongo")
```

## Curating the dataset with Ashy Drongo

```{r}
# add ashy drongo back to the BirdNET subset

# clean up column names
names(birdnet_subset)
names(birdnet_subset) <- c("start_time","end_time", "scientific_name","common_name","confidence_score","filepath","threshold","filename","date","site_name","site_protocol_code")
birdnet_subset <- birdnet_subset[,c(1:5,7:11)]

names(ashy)
names(ashy) <- c("selection","view","channel", "start_time","end_time", "low_freq", "high_freq", "filename_for_validation", "validity","confidence_score","site_name","date", "filename","site_protocol_code","scientific_name","common_name")
ashy <- ashy[,c(4,5,10:16)]
ashy$confidence_score <- as.numeric(ashy$confidence_score)

## create a combined dataset
datSubset <- bind_rows(birdnet_subset, ashy)

## create a unique detection id and separate out time
datSubset <- datSubset %>%
  mutate(
    # Extract time from filename (6 digits after the date)
    time = str_extract(filename, "\\d{6}(?=\\.WAV)"),
    
    # Remove .WAV and create unique detection ID with start and end times
    unique_detection_id = paste0(
      confidence_score, "_",
      site_name, "_",
      format(date, "%Y%m%d"), "_",
      time, "_",
      sprintf("%.1fs", start_time), "_",
      sprintf("%.1fs", end_time)
    )
  )

## write to file
write.csv(datSubset, "results/thresholded-acoustic-detections.csv", row.names = F)
```


